---
title: "Black First Achievements Timeline"
author: "Joanna Prout"
date: "2/21/2021"
output: html_document
---

This is my attempt at making a data viz from a challenge from @DiversityinData for Black History Month.  The data set from blackpast.org (available at https://data.world/diversityindata/february-2021-black-history-month) included a list of notable Black "first achievements" during the years 1738 - 2021 by category, along with the following quotes:


The 'first' achievements in various fields have historically marked footholds, often leading to more widespread cultural change. The shorthand phrase for this is "breaking the color barrier"

"[Black] history is about much more than chronicling a series of “firsts.”  The time and place of a breakthrough reflects not only remarkable individual achievement but is itself an indication of the progress or lack of progress of black people in realizing the centuries-old intertwined goals of freedom, equality, and justice." ~ blackpast.org

Datasource: Wikipedia & Kaggle


I wanted to capture both individual and collective impact over time.


```{r warning=FALSE, echo=TRUE, message=FALSE}
#Notable Black Achievements
library (tidyverse)
library (showtext)
library (patchwork)

df <- read.csv("https://query.data.world/s/conu5g63p5djt76z4v3x2liephcggn", 
               header=TRUE, stringsAsFactors=FALSE)

#color palette
bfly.col <- c("5F706A", "#D5B59B", "#C84F36", "#AE8B6D", "#2C2B2C", "#61A89D", "#E7965F", "#793324")

## Loading Google fonts (https://fonts.google.com/)
font_add_google("Bebas Neue")
font_add_google ("Lato")

showtext_auto ()

```

```{r data wrangling, echo = TRUE, warnng = FALSE, message = FALSE}
df1 <- df %>%
  group_by (year, category) %>%
  summarize (n = n()) %>%
  ungroup %>%
  group_by (category) %>%
  mutate (csum = cumsum (n))

allyears <- df1 %>% 
  group_by(category) %>% 
  complete(year = full_seq(1738:2021, 1)) %>% 
  group_by (category) %>%
  fill(csum)

allyears[is.na(allyears)] <- 0
allyears$category <- as.factor (allyears$category)

```



```{r stacked area chart, echo=TRUE}
plot <- ggplot (allyears, aes (x=year, y=csum, fill = category)) + 
  geom_area () + 
  scale_y_continuous (position = "right")

plot_total <- plot + theme (text = element_text (family = "Lato", size = 12),
                                   legend.position = "none",
                   panel.grid = element_blank (),
                   axis.line.x = element_line (),
                   axis.text.y = element_blank (),
                   axis.ticks.x = element_blank (),
                   axis.ticks.y = element_blank (),
                   panel.background = element_rect(color = "#E3E4E1", fill = "#E3E4E1"), #ice cube
                   plot.background = element_rect(color = "#e3e4e1", fill = "#e3e4e1")) +
            scale_fill_manual (values = bfly.col) +
  labs (x = "Year", y = "Achievements")

plot_total
```

```{r facetted, echo = TRUE}
facets <- ggplot (allyears, aes (x=year, y=csum, fill = category)) + 
    geom_area () +
    facet_wrap(~category, nrow = 2, strip.position = "bottom") 

facets

```

```{r facetted graph plus theme and colors, echo = TRUE}


facets_final <- facets +  
  labs (x = " ", y = " ", caption = "Data from blackpast.org | for @DiversityinData | graphic @joieprout", title = "\nBlack history is about more than chronicling a series of 'firsts'",
                  subtitle = "\nThe time and place of a breakthrough reflects \nnot only remarkable individual achievement\nbut is itself an indication of the progress \nor lack of progress of black people\nin realizing the centuries-old intertwined goals\nof freedom, equality, and justice. \n\n~ blackpast.org") +
  scale_fill_manual (values = c("5F706A", "#D5B59B", "#C84F36", "#AE8B6D", "#2C2B2C", "#61A89D", "#E7965F", "#793324")) +
  theme (text = element_text (family = "Lato", size = 15),
            plot.title = element_text (family = "Bebas Neue", face = "bold", size = 20),
                   plot.subtitle = element_text (family = "Lato", face = "italic", size = 12),
         legend.position = "none",
         strip.background = element_rect(
           color="#D5B59B", fill="#d5b59b", size=1.5, linetype="solid"
         ),
         panel.grid = element_blank (),
         panel.background = element_rect(color = "#e3e4e1", fill = "#e3e4e1"),
         plot.background = element_rect(color = "#e3e4e1", fill = "#e3e4e1"),
         axis.text.y = element_blank (),
         axis.ticks.y = element_blank (),
         axis.ticks.x = element_blank ()) 

facets_final
```


```{r combining plots using inset function, echo=TRUE}
combo <- facets_final + 
  inset_element(plot_total, left = 0.6, bottom = 0.5, right = 0.9, top = 0.8, align_to = 'full')

combo

```

```{r}
ggsave ("combo1.pdf", width = 10, height = 6)
```

